Empirical Bayes estimation of the normal inverse Gaussian prior in simultaneous-equation models

### 2019-05-22 docs/manuscript1.Rnw
commit before push

>>>>

  \section{General model 2}
  We have co-data $\mathbf{c}_{jd}$ available that describes the relationship
  between the covariate $j$ and outcome $d$. An example from eQTL is the 
  distance between SNP and gene. Model (\ref{eq:priormodel}) is then rewritten
  as:
  \begin{subequations}\label{eq:generalmodel2}
    \begin{align*}
      y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} 
      \(\x_i \tr \bm{\beta}_d, \sigma_d^2\), \nonumber \\
      \beta_{jd} | \gamma_{jd}^2, \sigma_d^2 & \sim \mathcal{N} 
      \(0, \sigma_d^2 \gamma_{jd}^2 \), \\
      \gamma_{jd}^2 & \sim \mathcal{IG}
      \( (\mathbf{c}_{jd} \tr \bm{\alpha})^{-1}, \lambda_{jd} \), \\
      \sigma_d^2 & \sim 1/\sigma_d^3, \nonumber
    \end{align*}
  \end{subequations}
  where now the variance components $\gamma_{jd}^2$ are both outcome and 
  covariate specific and modelled as a function of the co-data. 
  The formulation as in (\ref{eq:generalmodel2}) results in the following
  variational distribution for the $\gamma^2_{jd}$:
  $$
  q(\gamma_{jd}^2) \overset{D}{=} \mathcal{GIG}
  \(-1, \lambda_{jd}(\mathbf{c}_{jd} \tr \bm{\alpha})^2, \delta_{jd}\).
  $$
  The variational parameters are now updated by:
  \begin{align*}
    \bm{\Sigma}_d^{(h+1)} & = (a_d^{(h)})^{-1} 
    \[\X \tr \X + \diag(b_{jd}^{(h)})\]^{-1}, \\
    \bm{\mu}_d^{(h+1)} & = \[\X \tr \X + \diag(b_{jd}^{(h)})\]^{-1} 
    \X \tr \y_d, \\
    \delta_{jd}^{(h+1)} & = a_d^{(h)} \[(\bmu^{(h+1)}_{jd})^2 + 
    (\bSigma_d)_{jj}\] + \lambda_{jd}, \\ 
    \zeta_d^{(h+1)} & = \frac{1}{2} \bigg[\mathbf{y}_d \tr \mathbf{y}_d -
    2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + 
    \trace ( \X \tr \X \bm{\Sigma}_d^{(h+1)}) + 
    (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} \\
    & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + 
    \trace \[ \diag (b_{jd}^{(h)}) \bm{\Sigma}_d^{(h+1)}\] + 
    (\bm{\mu}_d^{(h+1)}) \tr \diag (b_{jd}^{(h)}) 
    \bm{\mu}_d^{(h+1)}\bigg],
  \end{align*}
  until convergence. The empirical Bayes step that consists of an iteratively 
  reweighted least squares regression of responses $(f_j^{(l)})^{-1}$ on
  predictors $\mathbf{z}_j$ with weights $\lambda_j^z f_j^{(l)}$, where 
  \begin{align*}
  f_j^{(l)} & = \E_{Q^{(h)}}(\tau_j) = [c_j^{(l)} - 
  (D + 1)/(\delta_j^z)^{(l)}] \cdot (\delta_j^z)^{(l)} [\mathbf{z}_j \tr 
  (\bm{\alpha}^z)^{(l)} ]^{-2} / (\lambda_j^z)^{(l)}, \\
  \lambda_j^z & = \[ c_j^{(l)} + f_j^{(l)} 
  (\mathbf{z}_j \tr \bm{\alpha}^z)^2 - 
    2\mathbf{z}_j \tr \bm{\alpha}^z \]^{-1}.
  \end{align*}

  Note that so far, the design matrix $\X$ is assumed to be shared across the
  equations. This is not a requirement, so that we may have $\X^d$, for 
  $d=1, \dots, D$ different design matrices. As a consequence, the index $j$
  may run over differently sized sequences.
